\documentclass[letterpaper,oneside,english,11pt,openany]{memoir}

\input{my-packages.tex}

\input{my-macros.tex}

\title{On the Mod \texorpdfstring{$p$}{p} Cohomology of Pro-\texorpdfstring{$p$}{p} Iwahori Subgroups}
% \date{\today}
\author{Daniel Kongsgaard}

\makeevenhead{ruled}{\scshape\leftmark}{}{\rightmark}
\makeoddhead{ruled}{\scshape\leftmark}{}{\rightmark}
\makeevenfoot{ruled}{}{\thepage}{}
\makeoddfoot{ruled}{}{\thepage}{}
\renewcommand{\footnotesize}{\fontsize{10}{12}\selectfont}
\setlength{\parindent}{0.5in}

\setulmarginsandblock{1.2in}{*}{0.8}
\setlrmarginsandblock{1in}{*}{1}
\setlength{\footskip}{\lowermargin}
\addtolength{\footskip}{-0.5in}
\checkandfixthelayout

\setsecnumdepth{subsection}
\settocdepth{subsection}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% BEGIN : DOCUMENT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%\OnehalfSpacing
\DoubleSpacing

% \maketitle

\hypersetup{pageanchor=false}
\pagestyle{empty}

\input{front.tex}

\clearpage
\null
\clearpage

\hypersetup{pageanchor=true}
\frontmatter

\pagestyle{plain}

\setcounter{page}{3}

\addcontentsline{toc}{chapter}{Dissertation Approval Page}

\noindent {\large The dissertation of Daniel Kongsgaard is approved, and it is acceptable in quality and form for publication on microfilm and electronically.}

\vfill

\begin{center}
  {\large University of California San Diego} \\[1em]
  {\large 2022}
\end{center}

\clearpage

\renewcommand*{\chaptitlefont}{\centering\Large\scshape}

\renewcommand{\contentsname}{Table of Contents}
\renewcommand{\cftchaptername}{\chaptername~}
\renewcommand{\cftappendixname}{\appendixname~}
\renewcommand*{\cftchapterleader}{\bfseries\cftdotfill{\cftchapterdotsep}}
\renewcommand*{\cftchapterdotsep}{\cftdotsep}
\tableofcontents

\clearpage

\printnomenclature[4cm]
% To update call: makeindex main.nlo -s nomencl.ist -o main.nls

\clearpage

\renewcommand{\cfttablename}{\tablename~}
\listoftables

\clearpage

\addcontentsline{toc}{chapter}{Acknowledgements}

\begin{center}
  \Large\scshape Acknowledgements
\end{center}

I would like to acknowledge Professor Claus Sørensen for giving me interesting number theory research topics, and for his input and ideas during my research. I also thank him for his feedback on early drafts of this dissertation.

I would also like to acknowledge Daniel M.\ Kane for introducing me to research problems in robust statistics and his input and later collaboration on this research. In addition I would like to thank my other collaborators Ilias Diakonikolas, Jerry Li, and Kevin Tian.

\clearpage

\addcontentsline{toc}{chapter}{Vita}


\begin{center}
  \Large\scshape Vita
\end{center}
\begin{description}
  \item[2016--2017] Teaching Assistant, Aarhus University
  \item[2017] Bachelor of Science in Mathematics, Aarhus University
  \item[2021] Candidate in Philosophy in Mathematics, University of California San Diego
  \item[2017--2022] Teaching Assistant, University of California San Diego
  \item[2022] Doctor of Philosophy in Mathematics, University of California San Diego
\end{description}

\begin{center}
  \Large\scshape Publications (all joint)
\end{center}
\begin{enumerate}[$\bullet$]
  \item Ilias Diakonikolas, Daniel Kane, and Daniel Kongsgaard. 2020. List-Decodable Mean Estimation via Iterative Multi-Filtering. In \emph{Advances in Neural Information Processing Systems}, H. Larochelle, M. Ranzato, R. Hadsell, M.F. Balcan and H. Lin (Eds.), Vol. 33. Curran Associates, Inc., 9312--9323.

  \item Ilias Diakonikolas, Daniel Kane, Daniel Kongsgaard, Jerry Li, and Kevin Tian. 2021. List-Decodable Mean Estimation in Nearly-PCA Time. In \emph{Advances in Neural Information Processing Systems}, M. Ranzato, A. Beygelzimer, Y. Dauphin, P.S. Liang, and J. Wortman Vaughan (Eds.), Vol. 34. Curran Associates, Inc., 10195--10208.

  \item Ilias Diakonikolas, Daniel Kane, Daniel Kongsgaard, Jerry Li, and Kevin Tian. 2021. Clustering Mixture Models in Almost-Linear Time via List-Decodable Mean Estimation. (Accepted for ACM Symposium on Theory of Computing (STOC 2022).)
\end{enumerate}

% \begin{center}
%   \Large\scshape Fields of study
% \end{center}

% \begin{vplace}[0.7]
%   Some extra stuff. Acknowledgements, vita, abstact.
% \end{vplace}

\clearpage

\addcontentsline{toc}{chapter}{Abstract of the Dissertation}
% \renewcommand{\abstractnamefont}{\normalfont\Large\scshape}
% \renewcommand{\abstractname}{Abstract of the Dissertation}
% \begin{abstract}
%   Let $\gs{G}$ be a split and connected reductive $\Z_{p}$-group and let $\gs{N}$ be the unipotent radical of a Borel subgroup. In the first chapter of this dissertation we study the cohomology with trivial $\F_{p}$-coefficients of the unipotent pro-$p$ group $N = \gs{N}(\Z_{p})$ and the Lie algebra $\lie{n} = \Lie(\gs{N}_{\F_{p}})$. We proceed by arguing that $N$ is a $p$-valued group using ideas of Schneider and Zábrádi, which by a result of Sørensen gives us a spectral sequence $E_{1}^{s,t} = H^{s,t}(\lie{g},\F_{p}) \Longrightarrow H^{s+t}(N,\F_{p})$, where $\lie{g} = \F_{p} \otimes_{\F_{p}[\pi]} \gr N$ is the graded $\F_{p}$-Lie algebra attached to $N$ as in Lazards work. We then argue that $\lie{g} \iso \lie{n}$ by looking at the Chevalley constants, and, using results of Polo and Tilouine and ideas from Große-Klönne, we show that the dimensions of the $\F_{p}$-cohomology of $\lie{n}$ and $N$ agree, which allows us to conclude that the spectral sequence collapses on the first page.

%   In the second chapter we study the mod $p$ cohomology of the pro-$p$ Iwahori subgroups $I$ of $\SL_{n}$ and $\GL_{n}$ over $\Q_{p}$ for $n=2,3,4$ and over a quadratic extension $F/\Q_{p}$ for $n=2$. Here we again use the spectral sequence $E_{1}^{s,t} = H^{s,t}(\lie{g},\F_{p}) \Longrightarrow H^{s+t}(I,\F_{p})$ due to Sørensen, but in this chapter we do explicit calculations with an ordered basis of $I$, which gives us a basis of $\lie{g} = \F_{p} \otimes_{\F_{p}[\pi]} I$ that we use to calculate $H^{s,t}(\lie{g},\F_{p})$. We note that the spectral sequence $E_{1}^{s,t} = H^{s,t}(\lie{g},\F_{p})$ collapses on the first page by noticing that all maps on each page are necessarily trivial. Finally we note some connections to cohomology of quaternion algebras over $\Q_{p}$ and point out some future research directions.
% \end{abstract}

\vspace*{1in}

\begin{center}
  {\Large\scshape Abstract of the Dissertation}\\[1.5em]
  {\large On the mod \texorpdfstring{$p$}{p} cohomology of pro-\texorpdfstring{$p$}{p} Iwahori subgroups}\\[1.5em]
  by \\[1.5em]
  Daniel Kongsgaard \\[1em]
  Doctor of Philosophy in Mathematics \\[1em]
  University of California San Diego \\[1.5em]
  Professor Claus Sørensen, Chair \\[2em]
\end{center}

Let $\gs{G}$ be a split and connected reductive $\Z_{p}$-group and let $\gs{N}$ be the unipotent radical of a Borel subgroup. In the first chapter of this dissertation we study the cohomology with trivial $\F_{p}$-coefficients of the unipotent pro-$p$ group $N = \gs{N}(\Z_{p})$ and the Lie algebra $\lie{n} = \Lie(\gs{N}_{\F_{p}})$. We proceed by arguing that $N$ is a $p$-valued group using ideas of Schneider and Zábrádi, which by a result of Sørensen gives us a spectral sequence $E_{1}^{s,t} = H^{s,t}(\lie{g},\F_{p}) \Longrightarrow H^{s+t}(N,\F_{p})$, where $\lie{g} = \F_{p} \otimes_{\F_{p}[\pi]} \gr N$ is the graded $\F_{p}$-Lie algebra attached to $N$ as in Lazards work. We then argue that $\lie{g} \iso \lie{n}$ by looking at the Chevalley constants, and, using results of Polo and Tilouine and ideas from Große-Klönne, we show that the dimensions of the $\F_{p}$-cohomology of $\lie{n}$ and $N$ agree, which allows us to conclude that the spectral sequence collapses on the first page.

In the second chapter we study the mod $p$ cohomology of the pro-$p$ Iwahori subgroups $I$ of $\SL_{n}$ and $\GL_{n}$ over $\Q_{p}$ for $n=2,3,4$ and over a quadratic extension $F/\Q_{p}$ for $n=2$. Here we again use the spectral sequence $E_{1}^{s,t} = H^{s,t}(\lie{g},\F_{p}) \Longrightarrow H^{s+t}(I,\F_{p})$ due to Sørensen, but in this chapter we do explicit calculations with an ordered basis of $I$, which gives us a basis of $\lie{g} = \F_{p} \otimes_{\F_{p}[\pi]} I$ that we use to calculate $H^{s,t}(\lie{g},\F_{p})$. We note that the spectral sequence $E_{1}^{s,t} = H^{s,t}(\lie{g},\F_{p})$ collapses on the first page by noticing that all maps on each page are necessarily trivial. Finally we note some connections to cohomology of quaternion algebras over $\Q_{p}$ and point out some future research directions.

\clearpage

\mainmatter

\renewcommand*{\chaptitlefont}{\Huge\bfseries}
%\pagestyle{ruled}

\input{Chapters/intro.tex}

\chapter{Cohomology of Unipotent Groups}%
\label{cha:cohunigps}

\input{Chapters/cohunigps.tex}

\chapter{Cohomology of pro-\texorpdfstring{$p$}{p} Iwahori Subgroups}%
\label{cha:cohiwagps}

\input{Chapters/cohiwagps.tex}

\clearpage

\appendix

\chapter{Calculations}

\input{Chapters/calculations.tex}

\chapter{Other Research}%
\label{cha:robstat}

\input{Chapters/robstat.tex}

\clearpage

\backmatter

% \pagestyle{plain}
\SingleSpacing

\printbibliography

\clearpage

\printindex

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% END : DOCUMENT %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
